### 2023-09-27 _analysis/PRT-Appendix-1.Rmd
rename _analysis

---
title: "PRT Appendix - 1"
date: "`r Sys.Date()`"
output: github_document
---

Table specifying phases, steps, and item inclusion with justifications from workshop and literature in support.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(tidyverse)
library(here)
memos <- readr::read_csv(here("./data/PRT-WorkshopCoding - Memo Manager.csv"))%>% 
  select(Name, Groups, Comment) %>% 
  mutate(phase = substr(Name, 1,1) %>%  as.numeric(),
         step = substr(Name, 3,3) %>% as.numeric())

#@TODO Alternative 2: Make new Groups in Atlas that match with the Phase and step names

#@TODO: for appendix, we want justifications and also quotations of sources.
#For the latter, we can filter out quotes that come from documents in the PDF group only, and then join those in.

codes <- readr::read_csv(here("data/PRT-WorkshopCoding - Code Manager.csv")) %>%  
  select(Name, Comment, Groups) %>% 
  mutate(phase = substr(Name, 1,1) %>%  as.numeric(),
         step = ifelse(str_detect(Groups, "step"), substr(Name, 3,3),"0") %>% as.numeric())

PRT_items <- bind_rows(codes, memos) %>% select(phase, step, Name, Comment) %>% arrange(phase, step)

knitr::kable(PRT_items) #@TODO how to nest

PRT_items <- 
  PRT_items %>% mutate(response = "")

```

```{r write-draft-template, echo = FALSE, eval = FALSE}
write.csv(PRT_items, file = here("./outputs/draft_template.csv"))
```

```{r session-info, echo = TRUE, eval = TRUE}
sessionInfo()
```