### 2023-09-27 _analysis/PreReg_May_2021.Rmd

---
title: "Ecology and Conservation Modelling Preregistration Template -- DRAFT"
author: "Chris Jones, Elliot Gould"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document: default
---

# 1. Problem Formulation

## 1.1 Model Context and Purpose

### 1.1.1. Key Stakeholders

*First, identify relevant interest groups:*
-   *Who is the model for?*
-   *Who is involved in formulating the model?*
-   *Who needs buy in?*
-   *How key stakeholders will/wish to be involved in model development and testing*

### 1.1.2 The problem, and subsequent purpose of the model
*Briefly outline:*

-   *The problem (including the decision to be made, if relevant).*
-   *How your model will address the problem, being clear about the scope i.e. is your model addressing the whole problem, or part of it? Are there any linked problems that your model needs to cognisant of?*
-   *The scope of the model, including temporal and spatial resolutions*
-   *Finish with a clear statement of the purpose of your model.*

### 1.1.3 Analytical purpose and scope, including scenario analysis

*Provide more detail on the analytical purpose and scope of the model:*
-   *How will the model be analysed?*
-   *What analytical questions will the model be used to answer?*
-   *What is the intended domain in which the model is to be applied? Are there any reasonable extrapolation limits beyond which you expect the model should not be applied? *
-   *Are there potentially unacceptable management or policy outcomes identified by stakeholders that should be captured in the model, i.e. as constraints?*
-   *Are there scenarios that model inputs or outputs must accommodate? Scenarios should be set a priori, i.e. ideally before the model is built..*
-   *If relevant, describe what processes you will use to elicit and identify relevant scenarios, e.g. literature review, structured workshops with stakeholders or decision-makers.*
-   *What degree of flexibility is required from the model? Might the model need to be quickly reconfigured to explore new scenarios or problems proposed by clients / managers / model-users?*
-   *Last, are there any limitations on model development analysis and flexibility, such as time or budget constraints, for example, does a model need to be deployed rapidly?*

### 1.1.4 Output (or response) variable(s)

-   *What variables would support taking a particular course of action (i.e. a decision), or help define the problem at hand?*
-   *What are the values to be addressed in this problem context? Are they all relevant to this model?*
- *After answering these questions with relevant stakeholders, please describe and define your response variable(s).*

### 1.1.5 Scenario analysis operationalization (delete as appropriate)

-   *How will you operationalize any scenarios identified in 1.1.3? For example, how will you operationalise any qualitative changes of interest, such as ‚ 'deterioration' or 'improvement'?*
-   *Describe how you will evaluate and distinguish the performance of alternative scenario outcomes. *
-   *Justify or otherwise explain how you chose these measures and determined performance criteria in relation to the analytical objectives, model purpose and modelling context, such as the risk attitudes of decision-makers and stakeholders within this system.*

# 2.0 Conceptual Model Definition

## 2.1 Conceptual Model Representation and Elicitation Method

-   *Describe what method you will use to elicit or identify the conceptual model.*
-   *How will you represent the final conceptual model? This will likely depend on the method chosen to elicit the conceptual model.*

### 2.2.1 Explain Critical Conceptual Design Decisions

*List and explain critical conceptual design decisions, including:*

-   *spatial and temporal scales,*
-   *selection of entities and processes,*
-   *representation of stochasticity and heterogeneity,*
-   *consideration of local versus global interactions, environmental drivers, etc.*
- *Explain and justify the influence of particular theories, concepts, or earlier models against alternative conceptual design decisions that might lead to alternative model structures .*

### 2.2.2 Model assumptions and uncertainties
*Specify key assumptions and uncertainties underlying the model design, describing how uncertainty and variation will be represented in the model. Sources of uncertainty may include:*

-   *exogenous uncertainties affecting the system,*
-   *parametric uncertainty in input data and*
-   *structural / conceptual nonparametric uncertainty in the model.*

## 2.3 Predictor Variables

*Identify and define system system variables and structures, referencing scenario definitions, and the scope of the model as described within problem formulation (1.1.3):*

-   *What variables would support taking this action or making this decision?*
-   *What additional variables may interact with this system (things we can't control, but can hopefully measure)?*
-   *What variables have not been measured, but may interact with the system (often occurs in field or observational studies)?*
-   *What variables are indice or surrogate measures of variables that we cannot or have not measured?*
-   *In what ways do we expect these variables to interact (model structures)?*

## 2.4 Define prior knowledge, data specification and evaluation

*This section should specify the plan for collecting, processing and preparing data available for parameterisation, determining model structure, and for scenario analysis.*

### 2.4.1 Document data sources

*Document the identity, quantity and provenance of any data that will be used to develop, identify and test the model. Describe how the data is arranged, in terms of replicates and covariates.*

**For pre-existing data (delete as appropriate):**

-   *For each dataset, is the data open or publicly available? (Y/N)*
-   *How can the data be accessed? Provide a link or contact as appropriate, indicating any restrictions on the use of data.*
-   *Date of download, access, or future access:*
-   *Describe the source of the data - what entity originally collected this data? (National Data Set, Private Organisational Data, Own Lab Collection, Other Lab Collection, External Contractor, Meta-Analysis, Expert Elicitation, Other).*
-   *Codebook and meta-data. If a Codebook or other meta-data is available, please link to it here and / or upload the document(s).*
-   *Prior work based on this dataset - Have you published / presented any previous work based on this dataset? Include any publications, conference presentations (papers, posters), or working papers (in-prep, unpublished, preprints) based ont this dataset you have worked on.*
-   *Unpublished Prior Research Activity - Describe any prior but unpublished research activity using these data. Be specific and transparent.*
-   *Prior knowledge of the current dataset - Describe any prior knowledge of or interaction with the dataset before commencing this study. For example, have you read any reports or publications about this data?*

**Sampling Plan (For data you will collect) – delete as appropriate:**

-   *Data collection procedures*
    -   *Please describe your data collection process, including how sites and transects or any other physical unit were selected and arranged.*
    -   *Describe any inclusion or exclusion rules, and the study timeline.*
-   *Sample Size: Describe the sample size of your study.*
-   *Sample Size Rationale: Describe how you determined the appropriate sample size for your study. It could include feasibility constraints, such as time, money or personnel.*
-   *If sample size cannot be specified, specify a stopping rule - i.e. how will you decide when to terminate your data collection?*

### 2.4.2 Data Processing and Preparation

*Describe any data preparation and processing steps, including manipulation of environmental layers (e.g. standardisation and geographic projection) or variable construction (e.g. Principal Component Analysis).*

### 2.4.3 Data Exploration or preliminary data analyses

*For each separate preliminary or investigatory analysis:*

-   *State what needs to be known to proceed with further decision-making about the modelling procedure, and why the analysis is necessary.*
-   *Explain how you will implement this analysis, as well as any techniques you will use to summarise and explore your data.*
-   *What method will you use to represent this analysis (graphical, tabular, or otherwise, describe)*
-   *Specify exactly which parts of the data will be used.*
-   *Describe how the results will be interpreted, listing each potential analytic decision, as well as the analysis finding that will trigger each decision, where possible.*

### 2.4.4 Data evaluation, exclusion, and missing data

-   *Describe how you will determine how reliable the data is for the given model purpose. Ideally, model input data should be internally consistent across temporal and spatial scales and resolutions, and appropriate to the problem at hand*
-   *Document any issues with data reliability.*
-   *How will you determine what data, if any, will be excluded from your analyses? How will outliers be handled? Describe rules for identifying outlier data, and for excluding a site, transect, quadrat, year or season, species, trait, etc.*
-   *How will you identify and deal with incomplete or missing data?*

## 2.5 Conceptual Model Evaluation

-   *Describe how your conceptual model will be critically evaluated. Evaluation includes both the completeness and suitability of the overall model structure.*
-   *How will you critically assess any simplifying assumptions?*
-   *Will the conceptual model evaluation process include consultation or feedback from a client, manager, or model user?*

# 3.0 Model Formalisation and Implementation

## 3.1 Model Class, Framework and Approach

-   *Describe what modelling framework, approach or class of model you will use to implement your model and relate your choice to the model purpose and analytical objectives described in 1.1.2 and 1.1.3.*

### 3.1.1 Operationalising Model Variables

-   *Specify how you will operationalise the response variables in the model. This should relate directly to the analytical and or management objectives specified during the problem formulation phase. Operationalisations could include:*
    -   *the extent of a response,*
    -   *an extreme value,*
    -   *a trend,*
    -   *a long-term mean,*
    -   *a probability distribution,*
    -   *a spatial pattern,*
    -   *a time-series,*
    -   *qualitative change, such as a direction of change or*
    -   *the frequency, location, or probability of some event occurring.*
-   *Provide a rationale for your choices, including why plausible alternatives were not chosen.*
-   *For the remaining model variables, specify each variable and their nature of their treatment (e.g. lumped/distributed, linear/non-linear, stochastic/deterministic, Jakeman, 2006).*

### 3.1.2 Model Structure

*Specify model structural features, including:*
-   *the functional form of interactions,*
-   *data structures,*
-   *measures used to specify links,*
-   *any bins or discretisation of continuous variables (Jakeman, 2006).*

### 3.1.3 Model Class / Family

-   *Specify which family of statistical distributions you will use in your model, and describe any transformations, or link functions.*
-   *Justify your decision based on the purpose, objectives, prior knowledge and logistical constraints specified in the problem formulation phase.*
-   *Include in your rationale for selection, detail about which variables the model outputs are sensitive to, what aspects of their behaviour are important, and any associated spatial or temporal dimensions in sampling.*

### 3.2 Model Structure and Parameter Estimation

-   *Specify what approach and methods you will use to identify model structure and parameters.*

### 3.2.1 Parameter/Structure estimation technique

-   *Specify what technique you will use to estimate parameter values, and how you will supply non-parametric variables and/or data (e.g. distributed boundary conditions). For example, will you calibrate all variables simultaneously by optimising fit of model outputs to observations, or will you parameterise the model in a piecemeal fashion by either direct measurement, inference from secondary data, or some combination (Jakeman, 2006).*
-   *Identify which variables were parameterised directly (if any), such as by expert elicitation or prior knowledge.*
-   *Specify which algorithm(s) you will use for any data-driven parameter estimation, including supervised, or unsupervised machine learning, decision-tree, K-nearest neighbour or cluster algorithms (Liu et al. 2018).*
-   *Repeat the above bullet points but for model structure estimation, if you are using quantitative methods for structure estimation rather than knowledge or expert-judgment driven methods.*

### 3.2.2 Estimation performance criteria

-   *Specify which suite of performance criteria you will use to judge the performance of the model. Examples include correlation scores, coefficient of determination, specificity, sensitivity, AUC, etcetera (Yates et al. 2018). State any numeric thresholds and their qualitative interpretation for judging adequate estimation.*
-   *Explain how you will identify which model features or components are significant or meaningful.*

## 3.3 Model assumptions and uncertainties

-   *Specify all assumptions and key uncertainties in the formal model.*
-   *Describe what gaps exist between the model conception, and the real-world problem, what biases might this introduce and how might this impact any interpretation of the model outputs, and*
- *How do these choices about what is included in the model and the way it has been operationalised in the formal model influence any inferences or conclusions you might draw? If the model is being used in a decision-making context, how might these assumptions influence what decisions you can make?*

## 3.4 Specify formal model(s)

-   *Once all decisions have been made about the approach and method of model specification, translate the conceptual model into the formal or quantitative model, i.e. provide mathematical, statistical equations, OR the model code for the models you will fit.*
-   *Some modelling approaches require you to specify initial starting values within the call to the software function, especially when using data-driven approaches to model structure definition and parameterisation, or when using formal model selection procedures -- describe any initial model specifications and parameterisations, including for any tune-in parameters.*

# 4.0 Model Calibration, Fitting & Checking

## 4.1 Model calibration and validation scheme

-   *Describe how you will validate and check the calibration of the model. The model may be tested on data independent of those used to parameterise the model (external validation), or the model may be cross-validated on random sub-samples of the data used to parameterise the model (internal cross-validation) .*
-   *Some models are sensitive to the hyper-parameters you select, e.g. min buckets etc., therefore model tuning using cross-validation needs to be done before the final model can be parameterised / estimated. Please describe any model tuning procedures for estimating hyper-parameters, if any.*
-   *Justify your choices*

### 4.1.2 Describe calibration/validation data

-   *If partitioning data for cross-validation:*
    -   *Describe the approach specifying the number of folds that will be created and the relative size of each.*
    -   *Describe how will you document and/or share the partitioned data such that the data partitioning and any subsequent modelling based on this partitioning can be computationally reproduced.*

-   *If using external / independent holdout data:*
    -   *Describe any known differences between the training and validation datasets and the size of the validation dataset.*
    -   *Describe how will you document and/or share the external data such that any subsequent modelling can be computationally reproduced.*
    -   *It is preferable that any independent data used for model evaluation remains unknown to modellers during the process of model building, please describe the relationship modellers have to model validation data, will independent datasets be known to any modeller or analyst involved in the model building process?*

## 4.2 Implementation verification

-   *What Quality Assurance measures will you take to verify the model has been correctly implemented as specified? Specifying up front quality assurance tests for implementation verification may help to avoid selective debugging and silent errors.*
    -   *Tests could include syntax checking of code, and code reviews by peers or in-house code-reviewers at the journal / publisher.*
    -   *Checks for verification implementation should include i) thoroughly checking for bugs or programming errors, and ii) whether the implemented model performs as dictated by the model description (Grimm et al. 2014).*
    -   *Technical measures include using unit tests, or in-built checks within functions to prevent potential errors.*

## 4.3 Model Checking

*"Model Checking" goes by many names ("calibration testing or validation", "conditional verification", "quantitative verification", "model output verification"), and refers to a series of analyses that assess a model's performance in representing the system of interest (Conn et al. 2018). Model checking therefore helps diagnose assumption violations, and reveals where a model might need to be altered to better represent the data, and therefore system (Conn et al. 2018). Quantitative model checking diagnostics include goodness of fit, tests on residuals or errors, such as for heteroscedascity, cross-correlation, and autocorrelation (Jakeman, Letcher, and Norton 2006).*

### 4.3.1 Quantitative model checking

*During this process, observed data, or data and patterns that guided model design and calibration is compared to model output in order to identify if and where there are any systematic differences.*

-   *Specify any diagnostics or tests you will use during model checking to assess a model's performance in representing the system of interest.*
-   *For each test, specify the criteria that you will use to interpret the outcome of the test in assessing the model's ability to sufficiently represent the gathered data used to develop and parameterise the model.*

### 4.3.2 Qualitative model checking

-   *Explain how you will qualitatively check the model, and whether and how you will include users and clients in the process. Face validation could include structured walk-throughs, or presenting descriptions, visualisations or summaries of model results to experts for assessment.*

### 4.3.3 Assumption violation checks

-   *Explain how you will demonstrate robustness to model assumptions and check for violations of model assumptions.*
-   *If you cannot perform quantitative assumption checks, describe what theoretical justifications would justify a lack of violation of or robustness to model assumptions.*
-   *If you cannot demonstrate or theoretically justify violation or robustness to assumptions, explain why not, and specify whether you will discuss assumption violations and their consequences for interpretation of model outputs.*
-  *If assumption violations cannot be avoided, explain how you will explore the consequences of assumption violations on the interpretation of results (To be completed in interim iterations of the preregistration, only if there are departures from assumptions as demonstrated in the planned tests above).*

# 5.0 Model Validation and Evaluation

## 5.1 Model output corroboration

-   *State whether you will corroborate the model outputs on external test data, and ensure that any independent validation data is documented in 4.1.2.*
-   *Testing the model against data that are collected spatially and/or temporally independent from the training dataset minimizes the risk of artificially inflating model performance measures (Araújo et al. 2019). Although the corroboration of model outputs against an independent validation dataset is considered the 'gold standard' for showing that a model properly represents the internal organisation of the system), model validation is not always possible (Grimm, et al. 2014). If you are unable to evaluate the model outputs against independent data, explain why not, and describe what steps you will take to interrogate the model. For example, Independent predictions can instead be tested on submodels, alternatively, patterns in model output that are robust and seem characteristic of the system can be identified and evaluated in consultation with the literature or by experts to judge how accurate the model's output is (Grimm et al.2014).*

### 5.1.1 Model corroboration performance metrics and criteria

-   *Specify what performance measures you will use to evaluate the model and briefly explain how each test relates to different desired properties of a model's performance. Examples include measures of agreement between predictions and independent observations, or estimates of accuracy, bias, calibration, discrimination refinement, resolution and skill (Araújo et al. 2019).*
-   *Spatial, temporal and environmental pattern of errors and variance can change the interpretation of model predictions and conservation decisions (Araújo et al. 2019), where relevant and possible, describe how you will characterise and report the spatial, temporal and environmental pattern of errors and variance.*
-   *If comparing alternative models, specify what measures of model comparison or out-of-sample performance metrics will you use to find support for alternative models or else to optimise predictive ability.*
-   *State what numerical threshold or qualities you will use for each of these metrics.*

## 5.2 Model Analysis

### 5.2.1 Uncertainty Analyses

-   *Please describe how you will characterise model and data uncertainties. E.g. propagating known uncertainties through the model, investigating the effect of different model scenarios with different parameters and modelling technique combinations (Araújo et al. 2019), or empirically constructing model distributions from model output data points, and extracting summary statistics, including the mean, median, variance, and constructing confidence intervals (Saltelli et al. 2019).*
-   *Relate your choice of analysis to the context and purposes of the model described in the problem formulation phase. For instance, high model precision may be important for predictive modelling when cost, benefit and risk must be adequately gauged, but for decision-making contexts where only the predicted ranked order of management options is necessary, model accuracy and precision might be less important (Jakeman, Letcher, and Norton 2006).*
-   *Describe how you will summarise the results of these in silico experiments with graphical, tabular, or other devices, such as summary statistics.*
-   *If the chosen modelling approach is able to explicitly articulate uncertainty due to data, measurements or baseline conditions, such as by providing estimates of uncertainty (typically in the form of probabilistic parameter covariance, (Jakeman, Letcher, and Norton 2006)), specify which measure of uncertainty you will measure.*

### 5.2.2 Sensitivity Analyses

*Sensitivity analysis examines how uncertainty in model outputs can be apportioned to different sources of uncertainty in model input (Saltelli et al. 2019).*

-   *Describe the sensitivity analysis approach you will take: deterministic sensitivity, stochastic sensitivity (variability in the model), or scenario sensitivity (effect of changes based on scenarios).*
-   *Describe any sensitivity analyses you will conduct by specifying which parameters will be held constant, which will be varied, and the range and intervals of values over which those parameters will be varied.*
-   *State the primary objective of each sensitivity analysis, for example, to identify which input variables contribute the most to model uncertainty so that these variables can be targeted for further data collection, or alternatively to identify which variables or factors contribute little to overall model outputs, and so can be 'dropped' from future iterations of the model (Saltelli et al. 2019).*

### 5.2.3 Model application / scenario analysis

-   *Specify any input conditions and relevant parameter values for initial environmental conditions and decision-variables under each scenario specified in section 1.*
-   *Please describe any other relevant technical details of model application, such as methods for how you will implement any simulations or model projections.*
-   *What raw and transformed model outputs will you extract from the model simulations or projections, and how will you map, plot, or otherwise display and synthesise the results of scenario and model analyses.*
-   *Explain how you will analyse the analysis output to answer your analytical objectives. For instance, describe any trade-off or robustness analyses you will undertake to help evaluate and choose between different alternatives in consultation with experts or decision-makers.*

### 5.2.4 Other in silico analyses

-   *Describe any other simulation experiments and or robustness analyses you will perform on the model, including any metrics and their criteria / thresholds for interpreting the results of the analysis.*

# References

Araújo, MB, RP Anderson, A Márcia Barbosa, CM Beale, CF Dormann, R Early, RA Garcia, et al. 2019. "Standards for Distribution Models in Biodiversity Assessments." Sci Adv 5 (1): eaat4858.

Augusiak, Jacqueline, Paul J Van den Brink, and Volker Grimm. 2014. "Merging Validation and Evaluation of Ecological Models to 'Evaludation': A Review of Terminology and a Practical Approach." Ecological Modelling 280: 117–28.

Barnard, David M., Matthew J. Germino, David S. Pilliod, Robert S. Arkle, Cara Applestein, Bill E. Davidson, and Matthew R. Fisk. 2019. "Can't See the Random Forest for the Decision Trees: Selecting Predictive Models for Restoration Ecology." Restoration Ecology.

Boets, Pieter, Dries Landuyt, Gert Everaert, Steven Broekx, and Peter L M Goethals. 2015. "Evaluation and Comparison of Data-Driven and Knowledge-Supported Bayesian Belief Networks to Assess the Habitat Suitability for Alien Macroinvertebrates" 74: 92–103.

Conn, Paul B, Devin S Johnson, Perry J Williams, Sharon R Melin, and Mevin B Hooten. 2018. "A Guide to Bayesian Model Checking for Ecologists." Ecological Monographs 9: 341–17.

Fraser, Hannah, Libby Rumpff, Jian D L Yen, Doug Robinson, and Brendan A Wintle. 2017. "Integrated Models to Support Multiobjective Ecological Restoration Decisions." Conservation Biology 31 (6): 1418–27.

Giljohann, K. M., McCarthy, M. A., McCarthy, M. A., Kelly, L. T., & Kelly, L. T. (2014). Choice of biodiversity index drives optimal fire management decisions. Ecological Modelling.

Gregory, R., Failing, L., Failing, L., Harstone, M., Harstone, M., Long, G., . . . Ohlson, D. (2012). Structured decision making: a practical guide to environmental management choices. John Wiley & Sons.

Grimm, Volker, Jacqueline Augusiak, Andreas Focks, Béatrice M Frank, Faten Gabsi, Alice S A Johnston, Chun Liu, et al. 2014. "Towards Better Modelling and Decision Support: Documenting Model Development, Testing, and Analysis Using Trace." Ecological Modelling 280: 129–39.

Jakeman, A J, R A Letcher, and J P Norton. 2006. "Ten Iterative Steps in Development and Evaluation of Environmental Models." Environmental Modelling & Software 21 (5): 602–14.

Liu, C. C., & Aitkin, M. (2008). Bayes factors: Prior sensitivity and model generalizability. Journal of Mathematical Psychology, 52(6), 362-375. <doi:10.1016/j.jmp.2008.03.002>

Liu, Zelin, Changhui Peng, Timothy Work, Jean-Noel Candau, Annie DesRochers, and Daniel Kneeshaw. 2018. "Application of Machine-Learning Methods in Forest Ecology: Recent Progress and Future Challenges." Environmental Reviews 26 (4): 339–50.

Mahmoud, Mohammed, Yuqiong Liu, Holly Hartmann, Steven Stewart, Thorsten Wagener, Darius Semmens, Robert Stewart, et al. 2009. "A Formal Framework for Scenario Development in Support of Environmental Decision-Making." Environmental Modelling & Software 24 (7): 798–808.

McDonald-Madden, Eve, Peter W. J. Baxter, and Hugh P. Possingham. 2008. "Making Robust Decisions for Conservation with Restricted Money and Knowledge." Journal of Applied Ecology 45 (6): 1630–8.

Moallemi, Enayat A., Sondoss Elsawah, and Michael J. Ryan. 2019. "Strengthening 'Good' Modelling Practices in Robust Decision Support: A Reporting Guideline for Combining Multiple Model-Based Methods." Mathematics and Computers in Simulation.

Saltelli, Andrea, Ksenia Aleksankina, William Becker, Pamela Fennell, Federico Ferretti, Niels Holst, Sushan Li, and Qiongli Wu. 2019. "Why so Many Published Sensitivity Analyses Are False: A Systematic Review of Sensitivity Analysis Practices." Environmental Modelling & Software 114: 29–39.

Tulloch, A. I. T., Maloney, R. F., Joseph, L. N., Bennett, J. R., Di Fonzo, M. M. I., Probert, W. J. M., . . . Possingham, H. P. (2015). Effect of risk aversion on prioritizing conservation projects. Conservation Biology.

White, Craig R, and Dustin J Marshall. 2019. "Should We Care If Models Are Phenomenological or Mechanistic." Trends in Ecology & Evolution 34 (4): 276–78.

Yates, KL, PJ Bouchet, MJ Caley, K Mengersen, CF Randin, S Parnell, AH Fielding, et al. 2018. "Outstanding Challenges in the Transferability of Ecological Models." Trends Ecol. Evol. (Amst.) 33 (10): 790–802.