### 2023-09-27 index.qmd
add quarto website scaffolding

---
title: "quarto"
---

This is a Quarto website.

To learn more about Quarto websites visit <https://quarto.org/docs/websites>.

```{r}
1 + 1
```