### 2023-09-27 RDF_QRP.qmd
Add scaffolding structure for website and incorp into _quarto.yml

---
title: From Researcher Degrees of Freedom to Questionable Research Practices
---