### 2023-09-27 adaptive_preregistration_background.qmd

---
title: A primer on Adaptive Preregistration
---