### 2023-09-27 adaptive_preregistration_implementation.qmd

---
title: Implementing Adaptive Preregistration with Git and Github
---