### 2023-09-27 RDF_QRP.qmd
Turn off sidebar

---
title: From Researcher Degrees of Freedom to Questionable Research Practices
sidebar: false
---