### 2023-09-27 preregistration_modelling.qmd
Add notes on adaptive preregistration

---
title: Adaptive Preregistration for Modelling
sidebar: false
---

How does adaptive preregistration differ from existing preregistration?

-   interim preregistrations for the iterative cycle of model development
-   registrered flexibility for when planned analyses need to change based on the outcomes of previous decisions
-   moving back and forth between different preregistration phases

Communicating deviations, additional analyses, outcomes of registered flexibility

-   fledge, News.md

Interim Preregistrations:

-   GitHub releases, semantic versioning, fledge