### 2025-03-03 _analysis/ms.qmd
Update Introduction & Methods

---
title: "But I can't preregister my research: Improving the reproducibility and transparency of ecology and conservation with adaptive preregistration for model-based research"
author:
  - id: aut1
    name:
      given: Elliot
      family: Gould
    email: elliot.gould@unimelb.edu.au
    orcid: 0000-0002-6585-538X
    attributes:
      corresponding: true
    roles:
      - conceptualization
      - data curation
      - formal analysis
      - funding acquisition
      - investigation
      - methodology
      - project administration
      - resources
      - software
      - validation
      - visualization
      - writing - original draft
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut2
    name:
      given: Chris
      family: Jones
    email: chris.jones@delwp.vic.gov.au
    orcid: 0000-0003-2833-0194
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - data curation
      - funding acquisition
      - investigation
      - methodology
      - project administration
      - resources
      - software
      - supervision
      - validation
      - visualization
      - writing - review & editing
    affiliations:
      - ref: aff2
  - id: aut3
    name:
      given: Jian
      family: Yen
    email: jian.yen@delwp.vic.gov.au
    orcid: 0000-0001-7964-923X
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - data curation
      - formal analysis
      - investigation
      - methodology
      - project administration
      - resources
      - software
      - supervision
      - validation
      - visualization
      - writing - review & editing
    affiliations:
      - ref: aff2
  - id: aut4
    name:
      given: Hannah
      family: Fraser
    email: hannahsfraser@gmail.com
    orcid: 0000-0003-2443-4463
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - investigation
      - methodology
      - supervision
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut5
    name:
      given: Henry
      family: Wootton
    email: henry.wootton@delwp.vic.gov.au
    orcid: 0000-0001-6506-0248
    attributes:
      corresponding: false
    roles:
      - data curation
      - formal analysis
      - investigation
      - methodology
      - resources
      - software
      - visualization
      - writing - review & editing
    affiliations:
      - ref: aff2
  - id: aut6
    name:
      given: Lyndsey
      family: Vivian
    email: lyndsey.vivian@delwp.vic.gov.au
    orcid: 0000-0003-1927-723X
    attributes:
      corresponding: false
    roles:
      - data curation
      - investigation
      - methodology
      - resources
      - writing - review & editing
    affiliations:
      - ref: aff2
  - id: aut7
    name:
      given: Megan
      family: Good
    email: goodm@unimelb.edu.au
    orcid: 0000-0002-6908-1633
    attributes:
      corresponding: false
    roles:
      - investigation
      - methodology
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut8
    name:
      given: David
      family: Duncan
    email: david.duncan@unimelb.edu.au
    orcid: 0000-0003-4411-8214
    attributes:
      corresponding: false
    roles:
      - investigation
      - methodology
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut9
    name:
      given: Libby
      family: Rumpff
    email: lrumpff@unimelb.edu.au
    orcid: 0000-0001-9400-8086
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - investigation
      - methodology
      - supervision
      - writing - review & editing
    affiliations:
      - ref: aff1
  - id: aut10
    name:
      given: Fiona
      family: Fidler
    email: fidlerfm@unimelb.edu.au
    orcid: 0000-0002-2700-2562
    attributes:
      corresponding: false
    roles:
      - conceptualization
      - funding acquisition
      - methodology
      - supervision
      - writing - review & editing
    affiliations:
      - ref: aff1
abstract: |
  1.	Preregistration is an open-science practice which aims to improve research transparency and mitigate questionable research practices, like cherry picking results. It helps protects against cognitive biases, like hindsight bias, that can influence how study outcomes are interpreted. There has been little uptake of preregistration in ecology and conservation, arguably because existing pre-registration templates focus on null-hypothesis significance testing whereas ecology and conservation often rely on different types of statistical modelling.
  2.	We argue that preregistration in model-based research in ecology and conservation is both possible and beneficial, using templates adapted for domain-specific methodologies. We applied a user-centred design approach to translate the concept of preregistration into model-based research practice for ecology and conservation.
  3.	We developed a generalised ‘modelling workflow’ to capture critical analysis decisions in the model development cycle, informing a preregistration template for model-based research. The iterative nature of model development is incompatible with existing, linear preregistration templates, creating a barrier to implementing preregistration for model-based research.
  4.	To better align the internal logic of preregistration with the iterative and non-linear process of ecological modelling, we propose, test and evaluate a methodology for ‘adaptive preregistration’, using a case-study of environmental flows modelling for decision-support in Victoria, Australia.
  5. This research provides a template and methodology for implementing adaptive preregistration of ecological models. Although we focus on ecology and conservation in this paper, the concept of adaptive preregistration, and the templates developed here, could be applied in model-based research in other scientific disciplines within science more broadly. Modellers need no longer cry “but I can’t preregister my research.”

keywords:
  - preregistration
  - transparency
  - reproducibility
  - modelling
  - ecology
  - conservation
  - open-science
  - research software engineering
affiliations:
  - id: aff1
    name: University of Melbourne
  - id: aff2
    name: Arthur Rylah Institute for Environmental Research
format:
  # preprint-typst:
    # fontsize: 11pt
    # mainfont: Libertinus Serif 
  # docx: default
  preprint-docx:
    toc: false
    number-sections: true
    highlight-style: github
  elsevier-pdf:
    journal:
      name: Journal Name
      formatting: preprint #review or doubleblind
      model: 3p
      layout: onecolumn
      cite-style: authoryear
      # graphical-abstract: "![](abstract.png)"
      # highlights:
        # - Highlight 1
        # - Highlight 2 
        # - Highlight 3
    keep-tex: true
bibliography: "references_obsidian.bib"
csl: methods-in-ecology-and-evolution.csl
---

# Introduction

Over the past ten years or so, a ‘replication crisis’ in some scientific fields has exposed problems with the reliability of research findings [@Neoh2023; @OpenScienceCollaboration:2015cn; @camererEvaluatingReplicabilitySocial2018; @Ioannidis:2014cm]. One such problem is the prevalence of Questionable Research Practices (QRPs), including p-hacking, Hypothesising After Results are Known (HARKing), and cherry-picking [@Wicherts2016]. QRPs increase the probability of false positive results, artificially inflate effect size estimates, and ultimately reduce the replicability of published findings, while encouraging overconfidence in the precision of results and exacerbating the risk of accepting and propagating false facts [@Hoffmann2021; @Roettger2019; @Nissen2016]. Self-reports from researchers show that QRPs are surprisingly common across disciplines [@John:2012eo; @Agnoli:2017kl; @Makel2023], including in ecology [@Fraser:2018cl; @kimmel2023]. Although these self-report surveys of QRPs are heavily focused on statistical significance testing research employing p-values, some broader concepts, like selective reporting/cherry picking, have counterparts in other inferential frameworks.

QRPs are precipitated when researchers opportunistically or systematically misuse ‘researcher degrees of freedom’ without disclosure [@Bakker2018; @Wicherts2016]. Researcher degrees of freedom refer to the many decisions and alternative analytic choices researchers make throughout the research process [@Simmons:2011iw; @Wicherts2016]. Analytic decisions are typically subjective but methodologically defensible, however, choices and omissions taken at each ‘fork’ may collectively and significantly influence the final ‘analysis strategy’ [*sensu* @Hoffmann2021], ‘specification’ [*sensu* @Simonsohn2020], or ‘modelling path’ [*sensu* @hamalainen2016]. The set of all (plausible or reasonable) possible decisions that a researcher could make over the entire analysis from beginning to end has been termed collectively as the ‘garden of forking paths’ [@Gelman2013] or the ‘multiverse’ [@Steegen2016]. ‘Many-studies’ have now demonstrated the consequences of the ‘garden of forking paths’, whereby alternative analytical decisions can lead researchers to make vastly different interpretations of results, with conflicting conclusions, even when investigating the same research question using the same underlying dataset [@Gould2025; @Hoogeveen2022; @Silberzahn2018; @Desbureaux2021; @Young2021].

Good intentions and simply being ‘aware’ of the potential for researcher degrees of freedom to lead to QRPs cannot sufficiently mitigate their risk because of the unconscious nature of their origin [@Nosek:2018eg; @Zvereva2021]. Fortunately, there are some potential solutions to this problem. Preregistration is an open-science practice that aims to distinguish between planned and ad hoc analyses [@Parker2019] thereby restricting opportunities for trialling analytical alternatives (‘researcher degrees of freedom’) and making any deviations from the preregistered plan transparent [@Wicherts2016]. Preregistration requires researchers to register their methods and analysis plan in a secure and publicly accessible platform prior to collecting and/or analysing data, which cannot be altered after submission [@Parker2019]. Preliminary empirical evidence indicates that studies with preregistered analysis plans are less likely to have positive or significant results and are more transparently reported [@brodeur2024; @toth2021], especially when the preregistration template includes detailed instructions and requirements [@Bakker2018] or the preregistered study is published as a ‘Registered Report’ [@scheel2021] – a new publication format that aims to combat selective reporting and publication bias wherein the peer-review process and decision to publish is made prior to the results are known.

One potential objection to preregistration for modelling is that model development is an intrinsically exploratory process [@McDermott2021], while preregistration is for confirmatory research [i.e. hypothesis or theory testing, @Prosperi2019]. These conceptions of modelling and preregistration are not entirely wrong, but they are oversimplified. In practice, the distinction between exploratory and confirmatory activities is not a sharp one [@Flournoy2020; @Devezer2020; @alspaugh2019]. Modelling includes some tasks and activities that are exploratory in nature, but it can be distinguished from wholly *exploratory analysis* in that modelling is often goal-driven and directed [@alspaugh2019]. In practice, researchers move back and forth between different types of reasoning and inference throughout the modelling process, with deductions leading to abductions as the analysis proceeds [@alspaugh2019; @Prosperi2019] or even occurring together in a hybridized manner [@Prosperi2019]. This bears a strong resemblance to more general philosophical observations about the iterative nature of science [@Chang2004].

In this iterative and non-linear process of modelling, initial analyses often inform future analyses of the same data whereby the modeller’s understanding of the problem and the model itself are incrementally adjusted in light of intermediate results from previous decision points [@hamalainen2016; @Schweinsberg2021; @Dwork2015]. Modellers are typically uncertain about which model is most appropriate for the data [@popovic2024] and are reluctant to risk registering a modelling strategy that may be incompatible with the data [@Roettger2019]. For example, modellers may perform exploratory analyses to validate distributional assumptions are met [@Campbell2021], which may result in subsequent changes to the model specification. Should data-dependent decisions like these be interpreted as violating a key tenet of preregistration, that is, failing to define the entire analysis protocol ahead of time [@Dwork2015; @Benning2019; @MacEachern2019]? Our answer to this is ‘no’, and within this paper we outline a methodology for implementing an expanded view of preregistration called *Adaptive Preregistration* [*sensu* @Srivastava2018], which captures decision points and reasoning about modelling choices, but which embraces the data-dependency practices that constitute the modelling development process. This goes beyond the work of others who offer discussion and guidance about when and how to deviate from a preregistration [@Lakens2024; @willroth2024], instead providing an approach to preregistration embodying the principles of *registered flexibility* [@Roettger2019; @Benning2019; @Srivastava2018] and iterative or *interim preregistrations* [@ioannidis2022; @hofman2023].

In this study we aim to translate preregistration for application to model-based research, specifically aimed at ecology, conservation and related disciplines. Here, we first identify a generalised ‘modelling workflow’ that captured critical analysis decisions in the model development cycle and use it to build an Adaptive Preregistration template that can work for ecological modelling. We then tested and evaluated our template using a case-study of environmental flows modelling for decision-support in Victoria, Australia. This is the first attempt at developing a fully-fledged preregistration template designed for model development and evaluation that we are aware of, and we propose it can be extended to work in disciplines outside of ecology.

# Methods

We modified a typical preregistration into model-based research contexts with a user-centred design framework, where users were engaged at all stages to produce an implementation of preregistration that is fit for purpose [@Pavelin2012, Fig. 1].

<!--# Insert Figure 1 Here -->

Figure 1 Process for translating preregistration to ecological modelling research [adapted from @Pavelin2012]. Note that the ‘Evaluate and Iterate’ component of the process was not undertaken in this study

## Problem Formulation: Defining Vision & Objectives for a preregistration template for model-based research

We first specified the research context, purpose, target users and use-cases, per (Pu et al. 2019a; Pavelin et al. 2012). Our context is model-based research, defined broadly as any research that uses quantitative modelling to answer its research question (Brudvig 2017). The purposes of the preregistration template and methodology are to:

1.  **Delimit 'researcher degrees of freedom'** [@pu2019a], to mitigate the the risk of questionable research practices when conducting model-based research in applied ecology and conservation.
2.  **Increase ‘research transparency’ of ecological modelling,** including both:
    1.  *production transparency* — including research artefacts like open-access data and materials, or data collection procedures [@Lupia2014],
    2.  *analytic transparency* — a complete account of how analytic conclusions are drawn from the data [@Lupia2014] where model choices, steps in the modelling process, assumptions and expectations about model outputs are clearly articulated [@Bodner2020].

We specified two use-cases for our Adaptive Preregistration template:

1.  Study authors or modellers seeking to preregister their modelling study,
2.  Editors, reviewers, and/or readers of the completed preregistration [@Pu2019; @pu2019a], who are tasked with verifying that analyses and findings reported in the final manuscript are consistent with the preregistered analysis plan.

Existing preregistration formats are designed primarily for study authors; they are static text-based documents that do not easily facilitate ‘preregistration checking.’ We designed our template to address this.

## Designing the General Modelling Workflow

We conducted a literature review to identify a modelling workflow suitable for application to a range of different modelling goals, e.g. exploration, inference and prediction, including both explanatory and anticipatory predictions [@Tredennick2021; @Mouquet2015; @Maris2018; @Shmueli:2010ec]; research contributions, e.g. development of new methods, or application / extension of existing methods and model transfers [@Yates2018]; types of models, e.g. phenomenological vs. mechanistic [@Connolly2017; @White2019; @lewis2023], across basic and applied contexts and across fields / disciplines in ecology and conservation [@Carmel2013; @Brudvig2017] while offering prescriptive guidelines based on ‘Good Modelling Practice’ [@Augusiak:2014gz]. The final output was a workflow reflecting a balance between idealised and actual modelling in practice.

We developed an Adaptive Preregistration methodology, so that users have the flexibility to adapt the preregistration template to their own needs and processes [see user guide <https://egouldo.github.io/EcoConsPreReg/>, @gould2024]. This can be facilitated in two ways: 1). The preregistration author can supply analytical or modelling heuristics or strategies that depend on the outcome of the previous decision-point or analysis, for example, decision-trees for certain points in the analysis or items in the preregistration template could be preregistered prior to analysing [@Baldwin2022a]. 2) using interim preregistrations that mark the transition between phases of modelling and analysis as different parts of the data are observed [@Srivastava2018].

Modellers can preregister ‘plans to deploy flexible strategies’ [@Srivastava2018] describing heuristics consisting of multiple analysis or modelling strategies, whose execution depends on the outcome of previous decision-points or analyses. For example, the modeller can preregister a ‘decision-tree’ [@Baldwin2022a] consisting of pre-defined rules about when a particular modelling strategy or decision should be implemented. Modellers can specify a ‘flexible’ strategy within their preregistration when faced with methodological or analytic uncertainty that cannot be resolved without observing parts of the data or by engaging in preliminary analyses, model checking, or other data-dependent decisions.

The modeller may follow an iterative process of preregistration in parallel with modelling, consisting of interim preregistrations that mark phases of modelling and analysis as different parts of the data are observe [@Srivastava2018]. As they proceed through the model development process the researcher shifts from ideation, preregistration and execution of the preregistered analysis plan and back again, generating interim preregistrations at different points within the model development process, depending on observed outcomes of the flexible heuristics described in the interim preregistration. The analyses executed in practice will depend on the observed outcomes of the registered flexible analyses.

Last, to facilitate transparency and preregistration checking of flexible modelling and analysis strategies, the results of any final and preliminary analyses must be linked back to the preregistered flexible strategy such that the link between a particular triggered analysis decision, and the analysis outcome that triggered the decision is explicit. Fig. 2 demonstrates a generalised process for using git and GitHub as the vehicle for transparently documenting Adaptive Preregistration of a modelling study. The procedure leverages [GitHub’s tag and release feature](https://help.github.com/en/github/administering-a-repository/releasing-projects-on-github) in conjunction with [Semantic Versioning](https://egouldo.github.io/EcoConsPreReg/(https:/semver.org/)) to track, document and collaborate on changes to the preregistration document. The preregistration is stored alongside modelling code and results with GitHub issues [see @braga2023 for explanation] tracking analysis tasks and results for a discrete task described in the preregistration and providing the opportunity to document interpretation of discussion informing the rationale for potential changes or updates to the preregistration that generate interim preregistrations. The use GitHub issues is particularly handy for registered flexible analyses: the outcome of the registered flexible analysis is recorded within that issue, linked to the described heuristic, and modellers can use the discussion features to document interpretation of those results, and provide interpretation to explain which registered decision alternative is to be selected in light of those results.

## Designing the Preregistration Template

We facilitated a workshop with 10 ecologists and ecological modellers. A key task was to identifycommon modelling workflows and critical decision-steps in model development. ‘Critical’ decision-steps are points in the modelling process where researchers, or modellers, make analytical decisions that could change the output of the analysis. In this activity, workshop participants first individually reflected on their own scientific process for a recent or memorable research project, listing the steps undertaken throughout the process. Next, participants collaboratively mapped the decision steps in their own personal modelling workflows onto a pre-prepared modelling workflow template (<https://osf.io/fgd23>), informed by our general workflow. We then collaboratively reviewed and refined the modelling workflow into a standard set of key modelling phases and commonly implemented decision steps.

<!--# Insert Figure 2 - Adaptive Preregistration Workflow -->

Figure 2 - Implementing Adaptive Preregistration. The preregistration (blue steps) and modelling and analysis processes (orange steps) operate in parallel. Purple tags represent git tags, corresponding to incremental versions of the preregistration.  Any changes to the analysis plan receive a GitHub release and major version number increment. Step 3 corresponds to the initial preregistration document, Step 6 corresponds to an interim preregistration, and Step 8 corresponds to the final preregistration. The execution of the Final Preregistration occurs in Step 9.